Constant-round Blind Classical Verification of Quantum Sampling

### 2020-04-15 qpip0.tex
Add/clarify some action items

Mahadev's end results in groundbreaking work 4 round protocol for BQP. At the core it is a 4 round protocol to let BQP prover commit XZ measurement to a classical machine.
}
Intuitively, the protocol allows the prover to commit to a state $\rho$ and the verifier to choose a measurement basis $h$ and learn a corresponding measurement outcome $M_{XZ}(\rho, h)$ \Ethan{Actually define this notation by cutting-pasting $h$ from verifier's input}.
\Ethan{Remove the next sentences or move to later}
Note that even a honest prover has complete freedom in choosing $\rho$;
what this protocol actually does is let the prover commit to a particular measurement on $\rho$.
This is good enough for our application, since we will feed the measurement results into the verification of $\QPIP_1$ protocol to ensure that the prover pick the history state as $\rho$ as we will show next.
We now describe the protocol while omitting the implementation details.

@@

Protocol:
\begin{enumerate}
	\item \label{step:measure1} The verifier generates a public and secret key pair $(pk, sk)\leftarrow\cVMeasure{1}(1^\lambda, h)$. It sends $pk$ to the prover.
	\item \label{step:measure2} The prover generates $(y, \sigma)\leftarrow\cPMeasure{2}(pk, \rho)$.
		$y$ is a classical ``commitment", and $\sigma$ is some internal state.
		He sends $y$ to the verifier.
	\item \label{step:measure3} The verifier samples $c\xleftarrow{\$}\zo$ uniformly at random and sends it to the prover. $c=0$ indicates what's called a ``testing round", and $c=1$ indicates a ``Hadamard round".
	\item \label{step:measure4} The prover generates a classical string $a\leftarrow\cPMeasure{4}(pk, c, \sigma)$ and sends it to back to the verifier.
	\item \label{step:output} If it's a testing round ($c=0$), then the verifier generates and outputs $o\leftarrow\cVMeasure{T}(pk, y, a)$, where $o\in\set{\Acc, \Rej}$. \Ethan{Just say testing and acc/rej}
		If it's \Ethan{No abbreviations} a Hadamard round ($c=1$), then the verifier generates and outputs $v\leftarrow\cVMeasure{H}(sk, h, y, a)$.
\end{enumerate}
\end{protocol}

@@

\begin{lemma}\label{lem:urmila-binding}[binding property of Protocol~\ref{proto:urmila4}\cite{FOCS:Mahadev18a}]
	Let $\PMeasureStar$ be a $\BQP$ cheating  prover for $\PiMeasure$ and $\lambda$ be the security parameter. Under the QLWE assumption, if $\PMeasureStar$ passes the testing round with probability $1-\negl(\lambda)$, then there exists some $\rho$ so that for all verifier's input $h \in \zo^n$, the verifier's outputs on the Hadamard round is computationally indistinguishable from $M_{XZ}(\rho, h)$.
\iffalse	
	Suppose that for all $\lambda\in\bbN$ and $h\in\zo^*$ \Ethan{or $\zo^n$?},
	$\PMeasureStar$ passes the testing round with probability $1-\negl(\lambda)$.
	Then, under the QLWE assumption, there exists some $\rho$ so that for all $h$,
	the verifier's outputs on the Hadamard round is $\negl(\lambda)$-computationally indistinguishable from $M_{XZ}(\rho, h)$.
\fi
\end{lemma}

@@

We now use $\PiMeasure$ to transform our $\QPIP_1$ Protocol for $\SampBQP$, $\PiSamp=(\PSamp, \VSamp)$, to a corresponding $\QPIP_0$ protocol $\PiNaive$ in a natural way.
Recall that in $\PiSamp$ the verifier learns the XZ measurement outcome of the prover's message.
In PiInt We let the verifier use $\PiMeasure$ to learn...

@@

	Protocol:
	\begin{enumerate}
		\item \label{step:naive1} The verifier chooses a $XZ$-measurement $h$ from the distribution specified in \stepref{qpip1-verify} of $\PiSamp$.
		\item \label{step:naive2} The prover prepares $\rho$ by running \stepref{qpip1-state-gen} of $\PiSamp$.
		\item \label{step:urmila-in-naive} The verifier and prover run $(\PMeasure(\rho), \VMeasure(h))(1^\lambda)$,
			where on $\PiMeasure$'s second message, the prover can instead choose to abort by sending $a=\bot$.
			This is useless for the cheating prover but useful for proving one of our theorems later. \Ethan{Where/how to say this?}
			If it's a testing round, the verifier accepts or rejects based on the outcome of $\PiMeasure$.
			If it's a Hadamard round, the verifier obtains $v$.
		\item \label{step:naive-output} If it's a Hadamard round, the verifier finishes the verification step of Protocol~\ref{ProtoQPIP1} by generating and outputting $(d, z)$

@@

\begin{lemma}\label{lem:naive-qpip0-binding}[binding property of $\PiNaive$]
	Let $\PNaiveStar$ be a cheating $\BQP$ prover for $\PiNaive$ and $\lambda$ be the security parameter.
	Under the QLWE assumption, if $\PNaiveStar$ passes the testing round with probability $1-\negl(\lambda)$,
	\Ethan{Wrong assumption. The ``perfect prover" notion here is generalized and allows giving up.}
	then the verifier's output in the Hadamard round is $(\varepsilon+\negl(\lambda))$-computationally indistinguishable from $(d, z_{ideal})$.
\end{lemma}

\begin{proof}
	\Ethan{TODO this is currently wrong and I'm not sure how to fix it.
	The version before this which had subnormalized states seems strange too.
	The correct proof should involve an alternative prover strategy that ``fixes" the abort case
	by re-running with perfect prover strategy,
	in order to connect the dots with \Cref{lem:urmila-binding}}
	By \Cref{lem:urmila-binding},
	there exist some $\rho$ such that
	$v\approx_{c,\negl(\lambda)}M_{XZ}(\rho,h)$.
	Let $\cVVer$ be the verification operation in \stepref{qpip1-verify} of $\PiSamp$.
	Since $\cVVer$ is efficient, we have
	$(d,z)\approx_{c, \negl(\lambda)}\cVVer(M_{XZ}(\rho, h))$.
	\Ethan{There were some shenanigans involving subnormalized states for which I'm not quite sure what's going on.}
	By \Cref{QPIP1thm}, for all $\rho$, 
	$\cVVer(M_{XZ}(\rho, h))\approx_{c, \varepsilon}(d, z_{ideal})$.
	So we conclude that
	$(d, z)\approx_{c, \varepsilon+\negl(\lambda)}(d, z_{ideal})$.
\end{proof}

@@

The following protocol is a $\QPIP_0$ protocol for $\SampBQP$

@@

Given inverse poly p(T), we can parameterize the protocol to have soundness p(T)

@@

\Ethan{TODO add discussion about parallel rep}

\Ethan{Define in words first, and use the protocol box to introduce notations

unfold for the sake of convenience in analysis

verifier h1 to hm, prover generates whatever it should generate (tensor of whatever),
pki ski expand all, use vec.

}

We now introduce our $\QPIP_0$ protocol $\PiSampZ$ for $\SampBQP$.
It is essentially a $m$-fold parallel repetition of $\PiNaive$.
However, instead of choosing whether to run test or Hadamard round randomly each round,
we run $m-1$ test rounds and only one Hadamard round.

@@

\Ethan{TODO Fix proof notations. Like U0, U notations might've been undefined; see parallel rep paper}

\begin{theorem}\label{thm:qpip0}
	$\PiSampZ$ is a $\QPIP_0$ protocol for $\SampBQP$ with $\negl(n)$ completeness error and computational soundness. 
	\Ethan{Its proof currently only shows $O(\varepsilon)$ accuracy.}
\end{theorem}

@@

%Let $S_m$ be sets of $\{c\}$ such that only one of the $c_i=1$.
\begin{lemma}[partition lemma (revision of Lemma 3.5 of \cite{parallelrep}\footnote{$G_{0}$ and $G_{1}$ of this version are created from doing $G$ of \cite{parallelrep} and post-select on the $th$ register being $0$ or $1$ then discard $ph,th,in$. Property~\ref{property:partition-err} corresponds to Property~1. Property~\ref{property:partition-testing}corresponds to Property~4, with $2^{m-1}$ changes to $m-1$ because we only have $m$ possible choices of $\{c\}$. Property~\ref{property:partition-binding} corresponds to Property~5. Property~\ref{property-partition-norm-sum} comes from the fact that $G_0$ and $G_1$ are post-selections of orthogonal results of the same $G$.})]\label{lem:partition2}
Let $(U_0,U)$ be a prover's strategy in Protocol~\ref{proto:QPIP0samp}, where $U_0$ is the how the prover generates answer in the second round, and $U$ is how the prover generates answer in the fourth round. Denote the string $0^{i-1}10^{m-i} \in \zo^m $ as $e_i$.  Let $\gamma_0 \in[0,1]$, and $T\in \mathbb{N}$ such that $\gamma_0=1/\poly(n)$ and $T=\poly(n)$.

 For all $i\in[m]$, $\gamma \in \L\{\frac{\gamma_0}{T},\frac{2\gamma_0}{T},\dots,\frac{T\gamma_0}{T}\R\}$, there exist two efficient quantum circuit $G_{0,i,\gamma}$ and $G_{1,i,\gamma}$ \Ethan{Not consistent with the cited paper; typeof(G0, G1) aren't q circuits} such that for all (possibly sub-normalized)  quantum state $\ket{\psi}_{\regX,\regZ}$,

@@

        \item \label{property:partition-testing}
For all $\{pk\}$, $\{y\}$, fixed $\gamma$, and  $j\neq i$, we have

@@

    % $\{c\}\in S_m$ such that $c_i = 0$
For all $\{pk\}$, $\{y\}$, fixed $\gamma$, and $j\neq i$, there exists an efficient quantum algorithm $\ext_i$ such that

@@

Now we are ready to prove Theorem~\ref{thm:qpip0}
\begin{proof}
Completeness is trivial.

@@

$$P_{acc,i,z}.$$
Note that $P_{acc,i,z}$ also depends on $\{\pk\},\,\{y\}$ and  $(sk_i,h_i,y_i)$ since it includes the measurement the prover did before sending $a$,  verifier's checking on $(m-1)$ copies of testing rounds, and  the verifier's final computation from $(sk_i,h_i,y_i,a_i)$.

>>>>

Mahadev's end results in groundbreaking work 4 round protocol for BQP. At the core it is a 4 round protocol to let BQP prover commit XZ measurement to a classical machine.
}
Intuitively, the protocol allows the prover to commit to a state $\rho$ and the verifier to choose a measurement basis $h$ and learn a corresponding measurement outcome $M_{XZ}(\rho, h)$ \Ethan{Actually define this notation by cutting-pasting $h$ from verifier's input}.
We now describe the protocol while omitting the implementation details.

@@

Protocol:
\begin{enumerate}
	\item \label{step:measure1} The verifier generates a public and secret key pair $(pk, sk)\leftarrow\cVMeasure{1}(1^\lambda, h)$. It sends $pk$ to the prover.
	\item \label{step:measure2} The prover generates $(y, \sigma)\leftarrow\cPMeasure{2}(pk, \rho)$.
		$y$ is a classical ``commitment", and $\sigma$ is some internal state.
		He sends $y$ to the verifier.
	\item \label{step:measure3} The verifier samples $c\xleftarrow{\$}\zo$ uniformly at random and sends it to the prover. $c=0$ indicates what's called a ``testing round", and $c=1$ indicates a ``Hadamard round".
	\item \label{step:measure4} The prover generates a classical string $a\leftarrow\cPMeasure{4}(pk, c, \sigma)$ and sends it to back to the verifier.
	\item \label{step:output} If it is a testing round ($c=0$), then the verifier generates and outputs $o\leftarrow\cVMeasure{T}(pk, y, a)$, where $o\in\set{\Acc, \Rej}$. \Ethan{Just say it's a test that can acc or rej}
		If it is a Hadamard round ($c=1$), then the verifier generates and outputs $v\leftarrow\cVMeasure{H}(sk, h, y, a)$.
\end{enumerate}
\end{protocol}

@@

\begin{lemma}[binding property of $\PiMeasure$]
\label{lem:urmila-binding}
	Let $\PMeasureStar$ be a $\BQP$ cheating  prover for $\PiMeasure$ and $\lambda$ be the security parameter. Under the QLWE assumption, if $\PMeasureStar$ passes the testing round with probability $1-\negl(\lambda)$, then there exists some $\rho$ so that for all verifier's input $h \in \zo^n$, the verifier's outputs on the Hadamard round is computationally indistinguishable from $M_{XZ}(\rho, h)$.
\iffalse	
	Suppose that for all $\lambda\in\bbN$ and $h\in\zo^*$ \Ethan{or $\zo^n$?},
	$\PMeasureStar$ passes the testing round with probability $1-\negl(\lambda)$.
	Then, under the QLWE assumption, there exists some $\rho$ so that for all $h$,
	the verifier's outputs on the Hadamard round is $\negl(\lambda)$-computationally indistinguishable from $M_{XZ}(\rho, h)$.
\fi
\end{lemma}

@@

We now use $\PiMeasure$ to transform our $\QPIP_1$ Protocol for $\SampBQP$, $\PiSamp=(\PSamp, \VSamp)$, to a corresponding $\QPIP_0$ protocol $\PiNaive$.
Recall that in $\PiSamp$ the verifier takes $X$ and $Z$ measurements on the prover's message.
In $\PiNaive$ we let the verifier use $\PiMeasure$ to learn those measurement outcomes instead.

@@

	Protocol:
	\begin{enumerate}
		\item \label{step:naive1} The verifier chooses a $XZ$-measurement $h$ from the distribution specified in \stepref{qpip1-verify} of $\PiSamp$.
		\item \label{step:naive2} The prover prepares $\rho$ by running \stepref{qpip1-state-gen} of $\PiSamp$.
		\item \label{step:urmila-in-naive}
		    The verifier and prover run $(\PMeasure(\rho), \VMeasure(h))(1^\lambda)$. On $\PiMeasure$'s second message, we allow the prover to abort by sending $a=\bot$, which does not benefit cheating provers but is useful for analysis.

			If it's a testing round, the verifier accepts or rejects based on the outcome of $\PiMeasure$.
			If it's a Hadamard round, the verifier obtains $v$.
		\item \label{step:naive-output} If it's a Hadamard round, the verifier finishes the verification step of Protocol~\ref{ProtoQPIP1} by generating and outputting $(d, z)$

@@

\begin{lemma}[binding property of $\PiNaive$]
\label{lem:naive-qpip0-binding}
	Let $\PNaiveStar$ be a cheating $\BQP$ prover for $\PiNaive$ and $\lambda$ be the security parameter.
	Under the QLWE assumption, if $\PNaiveStar$ passes the testing round with probability $1-\negl(\lambda)$,
	\Ethan{Wrong assumption. The ``perfect prover" notion here is generalized and allows giving up.}
	then the verifier's output in the Hadamard round is $(\varepsilon+\negl(\lambda))$-computationally indistinguishable from $(d, z_{ideal})$.
\end{lemma}

\begin{proof}
	\Ethan{TODO this is currently wrong and I'm not sure how to fix it.
	The version before this which had subnormalized states seems strange too.
	The correct proof should involve an alternative prover strategy that ``fixes" the abort case
	by re-running with perfect prover strategy,
	in order to connect the dots with \Cref{lem:urmila-binding}}
	By \Cref{lem:urmila-binding},
	there exist some $\rho$ such that
	$v\approx_{c,\negl(\lambda)}M_{XZ}(\rho,h)$.
	Let $\cVVer$ be the verification operation in \stepref{qpip1-verify} of $\PiSamp$.
	Since $\cVVer$ is efficient, we have
	$(d,z)\approx_{c, \negl(\lambda)}\cVVer(M_{XZ}(\rho, h))$.
	\Ethan{There were some shenanigans involving subnormalized states for which I'm not quite sure what's going on.}
	By \Cref{QPIP1thm}, for all $\rho$, 
	$\cVVer(M_{XZ}(\rho, h))\approx_{c, \varepsilon}(d, z_{ideal})$.
	So we conclude that
	$(d, z)\approx_{c, \varepsilon+\negl(\lambda)}(d, z_{ideal})$.

	\Ethan{Original proof below.}

	By Lemma~\ref{lem:urmila-binding}, if the prover passes the testing round with probability $1-\negl(n)$, there exist a $\rho$ such that $\norm{z_{commit}- M_{XZ}(\rho,h)}_q \leq \negl(n)$. Note that the prover can choose to abort, so $z_{commit}$ and $\rho$ are subnormalized. Let $Ver$ be the verification operation run by Protocol~\ref{ProtoQPIP1} on the measurement outcome. \hannote{describe better} Since $Ver$ is an efficient process and $Ver(z_{commmit})+p_{abort}(\Rej,\bot)=(d,z_{naive})$, we have that

\begin{equation} \label{eq:z=verM}
    \norm{(d,z_{naive})-(Ver( M_{XZ}(\rho,h))+p_{abort}(\Rej,\bot))}_q \leq \negl(n).
\end{equation}
\hannote{actually true with subnormalize shit?}
 By Theorem~\ref{QPIP1thm}, for all $\rho$,
 \begin{equation}
     \norm{Ver(M_{XZ}(\rho,h))-(d,z_{ideal})} \leq O(\eps).
 \end{equation}
   Combined with Equation~\ref{eq:z=verM} and Lemma~\ref{lem:computational-triangle0}, and note that $\norm{(\Rej,\bot)-(d,z_{ideal})}=0$, we have that
   \begin{equation}
       \norm{(d,z_{naive})-(d,z_{ideal})}_q\leq O(\eps)
   \end{equation}
\end{proof}

@@

\iffalse
The following protocol is a $\QPIP_0$ protocol for $\SampBQP$

@@

Given inverse poly p(T), we can parameterize the protocol to have soundness p(T)
\fi

@@

We now introduce our $\QPIP_0$ protocol $\PiSampZ$ for $\SampBQP$.
It is essentially a $m$-fold parallel repetition of $\PiNaive$.
However, instead of choosing whether to run test or Hadamard round randomly each round,
we run $m-1$ test rounds and only one Hadamard round.
In the description of our protocol below, we expand out the uses of $\PiNaive$ and $\PiMeasure$ in order to introduce notations that we need in our analysis.

@@

\Ethan{TODO Fix proof notations. Like U0, U notations might've been undefined; copy from parallel rep paper if needed}

\begin{theorem}\label{thm:qpip0}
	$\PiSampZ$ is a $\QPIP_0$ protocol for $\SampBQP$ with $\negl(n)$ completeness error and computational soundness. 
\end{theorem}

@@

%Let $S_m$ be sets of $\{c\}$ such that only one of the $c_i=1$.
\begin{lemma}[partition lemma (revision of Lemma 3.5 of \cite{parallelrep}\footnote{$G_{0}$ and $G_{1}$ of this version are created from doing $G$ of \cite{parallelrep} and post-select on the $th$ register being $0$ or $1$ then discard $ph,th,in$. Property~\ref{property:partition-err} corresponds to Property~1. Property~\ref{property:partition-testing} corresponds to Property~4, with $2^{m-1}$ changes to $m-1$ because we only have $m$ possible choices of $\{c\}$. Property~\ref{property:partition-binding} corresponds to Property~5. Property~\ref{property-partition-norm-sum} comes from the fact that $G_0$ and $G_1$ are post-selections of orthogonal results of the same $G$.})]\label{lem:partition2}
Let $(U_0,U)$ be a prover's strategy in Protocol~\ref{proto:QPIP0samp}, where $U_0$ is the how the prover generates answer in the second round, and $U$ is how the prover generates answer in the fourth round. Denote the string $0^{i-1}10^{m-i} \in \zo^m $ as $e_i$.  Let $\gamma_0 \in[0,1]$, and $T\in \mathbb{N}$ such that $\gamma_0=1/\poly(n)$ and $T=\poly(n)$.

 For all $i\in[m]$, $\gamma \in \L\{\frac{\gamma_0}{T},\frac{2\gamma_0}{T},\dots,\frac{T\gamma_0}{T}\R\}$, there exist two efficient quantum circuit $G_{0,i,\gamma}$ and $G_{1,i,\gamma}$ such that for all (possibly sub-normalized)  quantum state $\ket{\psi}_{\regX,\regZ}$,

@@

        \item \label{property:partition-testing}
For all $\vec{pk}$, $\vec{y}$, fixed $\gamma$, and  $j\neq i$, we have

@@

    % $\{c\}\in S_m$ such that $c_i = 0$
For all $\vec{pk}$, $\vec{y}$, fixed $\gamma$, and $j\neq i$, there exists an efficient quantum algorithm $\ext_i$ such that

@@

Now we are ready to prove Theorem~\ref{thm:qpip0}
\begin{proof}

\Ethan{This proof currently only shows $O(\varepsilon)$ accuracy, which doesn't perfectly match up with the definition. Hopefully doesn't matter way too much though?}

Completeness is trivial.

@@

$$P_{acc,i,z}.$$
Note that $P_{acc,i,z}$ also depends on $\vec{pk}, \vec{y}$, and $(sk_i, h_i)$ since it includes the measurement the prover did before sending $a$,  verifier's checking on $(m-1)$ copies of testing rounds, and  the verifier's final computation from $(sk_i,h_i,y_i,a_i)$.
